% In this file you should put the actual content of the blueprint.
% It will be used both by the web and the print version.
% It should *not* include the \begin{document}
%
% If you want to split the blueprint content into several files then
% the current file can be a simple sequence of \input. Otherwise It
% can start with a \section or \chapter for instance.

We synthesise material from \cite{Bochnak1998}, \cite{Knebusch2022}, \cite{Jacobson2012I} and \cite{Jacobson2012II} as convenient. 

Note that the meaning of the notation $F(\sqrt{a})$ depends on whether $a$ is a square in $F$. We take care to write it only in the non-trivial case. The same applies to $F(i)$.

\section{Ordered Fields} 

We begin with a purely algebraic characterisation of an ordered field. This relies on the theory of \textit{ring orderings}, which can be found in \cite{Lam1984}.

\begin{theorem}
  \label{thm:ord_iff_real}
  \lean{Field.exists_isStrictOrderedRing_iff_isSemireal}
  \leanok
  A field can be ordered if and only if it is real (that is, $-1$ is not a sum of squares).
\end{theorem}
\begin{proof}
  \leanok
\end{proof}

We can also characterise algebraically whether a field $F$ has a unique ordering. This allows us to talk about `the ordering' on $F$ without ambiguity.

\begin{lemma}
  \label{lem:unique_ord_cond}
  \lean{IsSemireal.existsUnique_isStrictOrderedRing_iff}
  \leanok
  Let $F$ be a real field. There is a unique ordering on $F$ if and only if, for each $a\in F$, either $a$ or $-a$ is a sum of squares. In this ordering, the non-negative elements are precisely the squares in $F$.
\end{lemma}
\begin{proof}
  \leanok
\end{proof}

\begin{corollary}
  \label{cor:unique_ord_cond_ordered}
  \lean{IsStrictOrderedRing.unique_isStrictOrderedRing_iff}
  \leanok
  Let $F$ be an ordered field. Then $F$ has a unique field ordering if and only if every non-negative element is a sum of squares.
\end{corollary}
\begin{proof}
  \uses{lem:unique_ord_cond}
  \leanok
\end{proof}

\begin{corollary}
  \label{cor:unique_ord_Q}
  \lean{Rat.unique_isStrictOrderedRing}
  \leanok
  There is a unique field ordering on $\Q$.
\end{corollary}
\begin{proof}
  \uses{cor:unique_ord_cond_ordered}
  \leanok
  We use Corollary \ref{cor:unique_ord_cond_ordered}. Let $x\in\Q$ be non-negative. Then $x=p/q$ for some integers $p\geq0$ and $q>0$, so 
  \[x=\underbrace{\frac{1}{q^2}+\cdots+\frac{1}{q^2}}_{pq\text{ times}}.\]
\end{proof}

There is a corresponding algebraic characterisation of an ordered field extension.

\begin{theorem}
  \label{thm:ext_ord_cond}
  \lean{Field.exists_isOrderedAlgebra_iff_neg_one_notMem_sup}
  \leanok
  Let $F$ be an ordered field, and let $K/F$ be a field extension. Then there is an ordering on $K$ making the extension $K/F$ ordered if and only if $\sum_i x_i\alpha_i^2\neq-1$ for all choices of $\alpha_i\in K$ and $x_i\in F_{\geq0}$.
\end{theorem}
\begin{proof}
  \leanok
\end{proof}

\begin{lemma}
  \label{lem:ext_ord_functional_suff}
  \lean{Field.exists_isOrderedAlgebra_of_projection}
  \leanok
  Let $F$ be an ordered field, and let $K/F$ be a field extension. Suppose that there is an $F$-linear functional $\pi:K\to F$ such that, for all $x\in K$, $\pi(x^2)\geq0$. Then there is a field ordering on $K$ making $K/F$ ordered.
\end{lemma}
\begin{proof}
  \uses{thm:ext_ord_cond}
  \leanok
  Consider the sum $\sum_i x_i\alpha_i^2$ for some $\alpha_i\in K$ and $x_i\in F_{\geq0}$. By $F$-linearity, we compute
  \[\pi\left(\sum_i x_i\alpha_i^2\right)=\sum_i x_i\pi(\alpha_i^2)\geq0.\]
  Since $\pi(-1)=-1<0$, we are done by Theorem \ref{thm:ext_ord_cond}.
\end{proof}

\begin{corollary}
  \label{cor:ext_ord_to_adj_sqrt}
  \lean{adj_sqrt_ordered}
  \leanok
  Let $F$ be a ordered field, and suppose $a\in F$ is non-negative (and not a square). Then there is a field ordering on $F(\sqrt{a})$ making $F(\sqrt{a})/F$ ordered.
\end{corollary}
\begin{proof}
  \uses{lem:ext_ord_functional_suff}
  \leanok
  Let $\pi:F(\sqrt{a})\to F$ be the projection induced by the $F$-basis $\{1,\sqrt{a}\}$. For $x,y\in F$, we have $\pi((x+y\sqrt{a})^2)=x^2+ay^2\geq0$. We are done by Lemma \ref{lem:ext_ord_functional_suff}.
\end{proof}

\begin{lemma}
  \label{lem:ext_ord_odd_deg}
  \lean{odd_deg_ordered}
  \leanok
  Let $F$ be an ordered field, and let $K/F$ be an odd-degree extension. Then there is a field ordering on $K$ making $K/F$ ordered.
\end{lemma}
\begin{proof}
  \uses{thm:ext_ord_cond}
  \leanok
  Since $\ch R=0$, we can apply the primitive element theorem. Let $K=F(\alpha)$ for some $\alpha\in K$, and let $f$ be the minimal polynomial of $\alpha$ over $K$. Then $\deg f=[K:F]$ is odd. By Theorem \ref{thm:ext_ord_cond}, we need to show that the congruence
  \[\sum_ia_ig_i^2\equiv-1\bmod f\tag{$\star$}\]
  fails to hold for any non-negative $a_i\in F$ and polynomials $g_i\in F[X]$ each of degree at most $\deg f-1$. Proceed by induction on $\deg f$; if $\deg f=1$, then ($\star$) reduces to an equality of a non-negative element of $F$ with a negative one. Otherwise, suppose for a contradiction that ($\star$) holds. Without loss of generality, we may assume, for all $i$, we have $a_i\neq0$ and that $\deg g_i<\deg f$.
  
  Rearranging ($\star$), we have $\sum_ia_ig_i^2+1=hf$ for some $h\in F[X]$. Let $d=\max_i\deg g_i$; note that $d<\deg f$ by construction. Since each $a_i$ is positive, the $2d$th coefficient on the left-hand side must be positive. Therefore
  \[\deg h+\deg f=\deg\left(\sum_ia_ig_i^2+1\right)=2d.\]
  Then $\deg h$ is odd, so $h$ has an odd-degree irreducible factor $\tilde h$. We have
  \[\deg\tilde h\leq\deg h=2d-\deg f<\deg f,\]
  but $\sum_ia_ig_i^2\equiv-1\bmod\tilde h$. We are done by induction.$_\#$
\end{proof}

There is an easier way to construct ordered field extensions if we don't care about them being algebraic.

\begin{lemma}
  \label{lem:order_fun_field}
  Let $F$ be an ordered field, and let $a\in F$. Then there is a unique ordering on the function field $F(X)$ making $F(X)/F$ ordered such that $X>a$ but $b>X$ for $b>a$, and a unique one such that $X<a$ but $b<X$ for $b<a$.
\end{lemma}

Intuitively, $X$ is infinitesimally close to $a$. When $a=0$, we often write $R(\eps)$ for $R(X)$ with the first type of ordering.

\section{Real Closed Fields}

\begin{definition}
  \label{def:RCF}
  \lean{IsRealClosed}
  \leanok
  A real closed field $F$ is a real field such that
  \begin{enumerate}
    \item for all $x$ in $F$, either $x$ or $-x$ is a square in $F$, and
    \item every odd-degree polynomial over $F$ has a root in $F$.
  \end{enumerate}
\end{definition}

\begin{lemma}
  \label{lem:RCF_ord_ax}
  \uses{def:RCF}
  \lean{IsRealClosed.of_linearOrderedField}
  \leanok
  An ordered field is real closed if
  \begin{enumerate}
    \item every non-negative element is a square, and
    \item every odd-degree polynomial over $F$ has a root in $F$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  \leanok
  Squares in an ordered field are non-negative.
\end{proof}

Fix a real closed field $R$.

\begin{lemma}
  \label{lem:RCF_sumsq_is_sq}
  \uses{def:RCF}
  \lean{IsSquare.of_isSumSq}
  \leanok
  Every sum of squares in $R$ is a square in $R$.
\end{lemma}
\begin{proof}
  \leanok
  If the negative of a sum of squares is a square, then, dividing through, $-1$ is a sum of squares, contradicting the realness of $R$.
\end{proof}

\begin{lemma}
  \label{lem:RCF_ord_unique}
  \uses{def:RCF}
  \lean{IsRealClosed.nonneg_iff_isSquare}
  \leanok
  $R$ has a unique ordering making it an ordered field. In this ordering, the non-negative elements are exactly the squares.
\end{lemma}
\begin{proof}
  \uses{lem:unique_ord_cond,lem:RCF_sumsq_is_sq}
  \leanok
  Follows directly from Lemmas \ref{lem:unique_ord_cond} and \ref{lem:RCF_sumsq_is_sq}.
\end{proof}

In what follows, all algebraic extensions are given up to $R$-isomorphism, as is conventional. Observe that, since $-1$ is not a square in $R$, $R(i)/R$ is a quadratic extension. We will show this is the \textbf{only} nontrivial algebraic extension of $R$.

\begin{lemma}
  \label{lem:alg_ext_odd_deg}
  \uses{def:RCF}
  \lean{IsRealClosed.odd_finrank_extension}
  \leanok
  There is no nontrivial odd-degree finite extension of $R$.
\end{lemma}
\begin{proof}
  \leanok
  Let $K/R$ be an odd-degree extension of $R$. By the primitive element theorem, $K=R(\alpha)$ for some $\alpha\in K$. Let $f$ be the minimal polynomial of $\alpha$ over $K$. Then $f$ is irreducible, but $\deg f=[K:R]$ is odd, so $f$ has a root in $R$. Therefore, $[K:R]=\deg f=1$; that is, $K\cong R$.
\end{proof}

\begin{lemma}
  \label{lem:ext_deg_2}
  \uses{def:RCF}
  \lean{IsRealClosed.isAdjoinRoot_i_of_isQuadraticExtension}
  \leanok
  The field $R(i)$ is the unique quadratic extension of $R$.
\end{lemma}
\begin{proof}
  \leanok
  Let $K/R$ be a quadratic extension. Since $\ch R\neq2$, we have $K\cong R(\sqrt{a})$ for some $a\in R$. Since $a$ cannot be a square in $R$, we know $-a$ must be one. Rescaling by $\sqrt{-a}$, we get $R(\sqrt{a})\cong R(i)$.
\end{proof}

\begin{lemma}
  \label{lem:Ri_ext_deg_2}
  \uses{def:RCF}
  \lean{IsRealClosed.finrank_neq_two_of_isAdjoinRoot_i}
  \leanok
  There is no quadratic extension of $R(i)$.
\end{lemma}
\begin{proof}
  \uses{lem:RCF_sumsq_is_sq}
  \leanok
  Since $\operatorname{char}R(i)\neq 2$, it suffices to show that every element of $R(i)$ is a square. Observe that every $x\in R$ is a square in $R(i)$: if $x$ is not a square in $R$, then $-x$ is, and so $x=(i\sqrt{-x})^2$. Now, fix $x=a+bi\in R(i)$ with $a,b\in R$. If $b=0$, then $a=x$ is a square in $R(i)$. Otherwise, we have $x=(c+di)^2$, where
  \[c=\sqrt{\frac{a+\sqrt{a^2+b^2}}{2}}\text{ and }d=\frac{b}{2c}.\]
  To see these are well-defined elements of $R(i)$, note that $\sqrt{a^2+b^2}$ is in $R$ by Lemma \ref{lem:RCF_sumsq_is_sq} and that $c\neq0$ because otherwise $-a=\sqrt{a^2+b^2}$ and so $b=0$.$_\#$
\end{proof}

\begin{theorem}
  \label{thm:FTAlg}
  \uses{def:RCF}
  \lean{IsRealClosed.finite_extension_classify}
  \leanok
  The only finite extensions of $R$ are $R$ itself and $R(i)$.
\end{theorem}
\begin{proof}
  \uses{lem:ext_deg_2, lem:Ri_ext_deg_2, lem:alg_ext_odd_deg}
  \leanok
  By separability, every finite extension of $R$ is contained in a finite Galois extension. Since $R(i)/R$ has no intermediate fields, it suffices to show the result for finite Galois extensions.
  
  Let $K/R$ be a nontrivial Galois extension of degree $2^k\cdot a$, where $k\geq 0$ and $a\geq1$ is odd. Applying the Galois correspondence to a Sylow 2-subgroup of $\Gal(K/R)$ yields an intermediate extension of degree $a$; by Lemma \ref{lem:alg_ext_odd_deg}, we have $a=1$ (and $k>0$). If $k>1$, iterating the last construction yields intermediate extensions $K/L/M/R$ with $[L:M]=[M:R]=2$. By Lemma \ref{lem:ext_deg_2}, $M\cong R(i)$, contradicting Lemma \ref{lem:Ri_ext_deg_2}.$_\#$ Therefore $k=1$ and (by Lemma \ref{lem:ext_deg_2}) $K\cong R(i)$.
\end{proof}

\begin{corollary}
  \label{cor:FTAlg_alg}
  \uses{def:RCF}
  \lean{IsRealClosed.instFiniteDimensionalOfIsAlgebraic}
  \leanok
  The only algebraic extensions of $R$ are $R$ itself and $R(i)$.
\end{corollary}
\begin{proof}
  \uses{thm:FTAlg}
  \leanok
  An infinite algebraic extension contains finite subextensions of arbitrarily large degree.
\end{proof}

\begin{corollary}
  \label{cor:RCF_ac}
  \uses{def:RCF}
  \lean{IsRealClosed.isAdjoinRoot_i_of_isAlgClosure}
  \leanok
  $\bar{R}=R(i)$.
\end{corollary}
\begin{proof}
  \uses{cor:FTAlg_alg}
  \leanok
\end{proof}

The converse to Theorem \ref{thm:FTAlg} is much easier.

\begin{lemma}
  \label{lem:sumsq_is_sq}
  \lean{isSquare_of_isSumSq_of_forall_adjoinRoot_i_isSquare}
  \leanok
  Let $F$ be a field in which $-1$ is not a square, and suppose every element of $F(i)$ is a square. Then sums of squares in $F$ are squares in $F$.
\end{lemma}
\begin{proof}
  \leanok
  Given $a,b\in F$, find $c,d\in F$ such that $a+bi=(c+di)^2$ in $F(i)$. Then $a=c^2-d^2$ and $b=2cd$, so $a^2+b^2=(c^2+d^2)^2$ is a square in $F$. By induction, every sum of squares in $F$ is a square in $F$.
\end{proof}

In particular, $F$ as in the last lemma is real.

\begin{lemma}
  \label{lem:FTAlg_converse}
  \uses{def:RCF}
  \lean{IsRealClosed.of_isAdjoinRoot_i_or_finrank_eq_one}
  \leanok
  Suppose $R$ is a field with unique nontrivial finite extension $R(i)$. Then $R$ is real closed.
\end{lemma}
\begin{proof}
  \uses{lem:sumsq_is_sq}
  \leanok
  Since $R(i)$ has no quadratic extensions, we can apply Lemma \ref{lem:sumsq_is_sq}: sums of squares in $R$ are squares in $R$, and $R$ is real.

  Take a non-square $a\in R$; then $R(\sqrt{a})\cong R(i)$. Suppose $i$ maps to $x+y\sqrt{a}$ for some $x,y\in R$. Then $-1=x^2+ay^2+2xy\sqrt{a}$; comparing coefficients, we get $-1=x^2+ay^2$. Since $-1$ is not a square, $y$ must be nonzero, and so $-a=(x/y)^2+(1/y)^2$ is a square.

  Now, fix a nonlinear odd-degree polynomial $f\in R[X]$. Then $R[X]/(f)$ cannot be a field since $R$ has no nontrivial odd-degree extensions, and so $f$ must be reducible. We are done by induction on the degree.
\end{proof}

\begin{corollary}
  \label{lem:FTAlg_converse_alg_closure}
  \uses{def:RCF}
  \lean{IsRealClosed.of_isAdjoinRoot_i_isAlgClosure}
  \leanok
  Suppose $R$ is a field with nontrivial algebraic closure $R(i)$. Then $R$ is real closed.
\end{corollary}
\begin{proof}
  \uses{lem:FTAlg_converse}
  \leanok
\end{proof}

As before, let $R$ be a real closed field. Theorem \ref{thm:FTAlg} is a powerful tool for deriving more of its properties.

\begin{lemma}
  \label{lem:RCF_max}
  \uses{def:RCF}
  \lean{IsRealClosed.maximal_isSemireal}
  \leanok
  $R$ has no nontrivial real algebraic extensions.
\end{lemma}
\begin{proof}
  \uses{cor:FTAlg_alg}
  \leanok
  The field $R(i)$ is not real since $-1$ is a square in it. We are done by Corollary \ref{cor:FTAlg_alg}.
\end{proof}

\begin{corollary}
  \label{lem:RCF_max_ord}
  \uses{def:RCF,lem:RCF_ord_unique}
  \lean{IsRealClosed.maximal_isSemireal}
  \leanok
  $R$ has no nontrivial ordered algebraic extensions (with respect to the unique order).
\end{corollary}
\begin{proof}
  \uses{lem:RCF_max,thm:ord_iff_real}
  \leanok
  Since ordered fields are real, we are done by Lemma \ref{lem:RCF_max}.
\end{proof}

The next property is a little less obvious.

\begin{lemma}
  \label{lem:irreds_class}
  \uses{def:RCF}
  \lean{IsRealClosed.irred_poly_classify}
  \leanok
  The monic irreducible polynomials over $R[X]$ have form $X-c$ for some $c\in R$ or $(X-a)^2+b^2$ for some $a,b\in R$ with $b\neq0$.
\end{lemma}
\begin{proof}
  \uses{thm:FTAlg}
  \leanok
  Let $f\in R[X]$ be monic and irreducible. The field $R_f=R[X]/(f)$ is a finite extension of $R$, so it is classified by Theorem \ref{thm:FTAlg}. If $R_f\cong R$, then $\deg f=1$, so $f=X-c$ for some $c\in R$. If $R_f\cong R(i)$, let the isomorphism be $\varphi$, and suppose $\varphi(X+(f))=a+bi$ ($a,b\in R$). Note that $b\neq0$ since $\varphi^{-1}$ is constant on $R$. Rearranging, we see that $\varphi((X-a)^2+b^2+(f))=0$; that is, $(X-a)^2+b^2\in(f)$. Since this polynomial is monic and has the same degree as $f$, it must in fact be equal to $f$.

  Conversely, linear polynomials over a domain are irreducible by degree, and reducible quadratics have a root. A root of $f=(X-a)^2+b^2$ with $a,b\in R$ is an element $r\in R$ satisfying $(r-a)^2=-b^2$. If $b\neq0$, then dividing through would give a contradiction to the realness of $R$.Therefore $f$ must be irreducible.
\end{proof}

\begin{lemma}
  \label{lem:IVP_poly}
  \uses{def:RCF, lem:RCF_ord_unique}
  \lean{IsRealClosed.intermediate_value_property}
  \leanok
  Polynomials over $R$ satisfy the intermediate value property (with respect to the unique order).
\end{lemma}
\begin{proof}
  \uses{lem:irreds_class}
  \leanok
  We will prove that, for all $f\in R[X]$ and all $a,b\in R$ with $a\leq b$, if $f(a)\leq 0\leq f(b)$, then there is some $c\in[a,b]$ such that $f(c)=0$.

  Fix $a,b\in R$ with $a\leq b$, and proceed by induction on $\deg f$. If $\deg f=0$, then $f$ is constant and the result is clear. Otherwise, take a monic irreducible factor $g$ of $f$; then $g$ is classified by Lemma \ref{lem:irreds_class}.

  Observe that, if $g(a)$ and $g(b)$ are both positive, then $f/g$ satisfies the inductive hypothesis, and so has a root in $[a,b]$. If $g=(X-a)^2+b^2$ with $a,b\in R$ and $b\neq0$, then $g$ is everywhere positive. If $g=X-c$ with $c\in R$, then either $c\in[a,b]$ and $g(c)=0$, or $c\notin[a,b]$ and $f(a)$ and $f(b)$ have the same sign. In this last case, either $f/g$ or $f/(-g)$ satsifies the inductive hypothesis. In all cases, $f$ has a root in $[a,b]$.
\end{proof}

In fact, the converses to Lemmas \ref{lem:RCF_max} and \ref{lem:IVP_poly} also hold!

\begin{theorem}
  \label{thm:IVP_poly_imp_RCF}
  \uses{def:RCF}
  \lean{IsRealClosed.of_intermediateValueProperty}
  \leanok
  Let $R$ be an ordered field whose polynomials satisfy the intermediate value property. Then $R$ is real closed.
\end{theorem}
\begin{proof}
  \uses{lem:RCF_ord_ax}
  \leanok
  We use Lemma \ref{lem:RCF_ord_ax}. Let $a\in R$ be non-negative, and consider the polynomial $f=X^2-a$. Then $f(0)=-a\leq0$, but $f(a+1)=a^2+a+1>0$. By the intermediate value property, $f$ has a root in $R$, and so $a$ is a square in $R$.

  Let $f$ be an odd-degree polynomial over $R$. Write $f=a_nX^n+\cdots+a_0$. We will show $f$ has a root in $R$. Replacing $f$ by $-f$ if necessary, we may assume $a_n>0$. For $x>1$, we compute
  \[f(x)\geq x^{n-1}(a_nx-n\max_i|a_i|).\]
  Therefore, when $x>\max\{1,n\max_i|a_i|/a_n\}$, $f(x)>0$. A similar calculation shows that $f(x)<0$ for sufficiently large negative values of $x$. We are done by the intermediate value property.
\end{proof}

\begin{theorem}
  \label{thm:ord_max_imp_RCF}
  \uses{def:RCF}
  \lean{IsRealClosed.of_maximal_isOrderedAlgebra}
  \leanok
  Let $R$ be an ordered field with no nontrivial ordered algebraic extensions. Then $R$ is real closed.
\end{theorem}
\begin{proof}
  \uses{lem:RCF_ord_ax,cor:ext_ord_to_adj_sqrt,lem:ext_ord_odd_deg}
  \leanok
  We use Lemma \ref{lem:RCF_ord_ax}. 

  Let $a\in R$ be non-negative, and suppose $a$ is not a square. By Corollary \ref{cor:ext_ord_to_adj_sqrt}, there is an ordering making the nontrivial extension $R(\sqrt{a})/R$ ordered.$_\#$ Therefore $a$ is a square in $R$.

  By induction on degree, it suffices to show that irreducible odd-degree polynomials over $R$ are all linear. Let $f\in R[X]$ be such a polynomial, and consider the odd-degree field extension $R_f=R[X]/(f)$. By Lemma \ref{lem:ext_ord_odd_deg}, there is an ordering making $R_f/R$ an ordered extension. Therefore $\deg f=[R_f:R]=1$.
\end{proof}

\begin{corollary}
  \label{cor:real_max_imp_RCF}
  \uses{def:RCF}
  \lean{IsRealClosed.of_maximal_isSemireal}
  \leanok
  Let $R$ be a real field with no nontrivial real algebraic extensions. Then $R$ is real closed.
\end{corollary}
\begin{proof}
  \uses{thm:ord_iff_real,thm:ord_max_imp_RCF}
  \leanok
  Since $R$ can be ordered and ordered fields are real, we are done by Theorem \ref{thm:ord_max_imp_RCF}.
\end{proof}

Theorem \ref{thm:ord_max_imp_RCF} gives us a way to ``construct'' real closed fields.

\begin{lemma}
  \label{lem:ACF_ind_2_RCF}
  \uses{def:RCF}
  An algebraically closed field of characteristic 0 has an index-2 real closed subfield.
\end{lemma}
\begin{proof}
  \uses{lem:order_fun_field, thm:ord_max_imp_RCF, cor:RCF_ac}
  Let $C$ be an algebraically closed field of characteristic 0. Observe that the prime subfield $\Q$ can be ordered. Further, given an ordered subfield $F$ with $\bar{F}\neq C$, we can use Lemma \ref{lem:order_fun_field} to adjoin an element transcendental over $F$, obtaining a strictly bigger ordered subfield.
  
  Apply Zorn's lemma to obtain a maximal ordered subfield $R\subseteq C$; then $\bar{R}=C$. By Theorem \ref{thm:ord_max_imp_RCF}, $R$ must be real closed. By Corollary \ref{cor:RCF_ac}, $C\cong R(i)$, and so $[C:R]=2$.
\end{proof}

In summary, we have the following characterisations of real closed fields.

\begin{theorem}
  \label{thm:RCF_tfae}
  \uses{def:RCF}
  \lean{IsRealClosed.TFAE}
  \leanok
  Let $R$ be a field. TFAE:
  \begin{enumerate}
    \item $R$ is real closed.
    \item $\bar{R}=R(i)$ (and $R(i)\neq R$).
    \item $R$ is real, but has no nontrivial real algebraic extensions.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \uses{cor:RCF_ac,lem:FTAlg_converse_alg_closure,lem:RCF_max,cor:real_max_imp_RCF}
  \leanok
\end{proof}

\begin{theorem}
  \label{thm:RCF_tfae_ord}
  \uses{def:RCF}
  \lean{IsRealClosed.TFAE_linearOrderedField}
  \leanok
  Let $R$ be an ordered field. TFAE:
  \begin{enumerate}
    \item $R$ is real closed.
    \item $R$ is maximal with respect to ordered algebraic extensions.
    \item Polynomials over $R$ satisfy the intermediate value property.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \uses{lem:IVP_poly,thm:IVP_poly_imp_RCF,lem:RCF_max_ord,thm:ord_max_imp_RCF}
  \leanok
\end{proof}

\section{Real Closures}

\begin{definition}
  \label{def:real_closure}
  \uses{def:RCF}
  \lean{IsRealClosure}
  \leanok
  Let $F$ be an ordered field. A real closure of $F$ is a real closed ordered algebraic extension of $F$.
\end{definition}

\begin{lemma}
  \label{lem:real_closure_exists}
  \uses{def:real_closure}
  Let $F$ be an ordered field. Then $F$ has a real closure.
\end{lemma}
\begin{proof}
  \uses{thm:ord_max_imp_RCF}
  Apply Zorn's lemma to ordered algebraic extensions of $F$. We are done by Theorem \ref{thm:ord_max_imp_RCF}.
\end{proof}

Just like with the algebraic closure, it makes sense to talk of \textit{the} real closure of an ordered field. Proving this uniqueness result requires a method of root-counting in real fields known as Sturm's theorem.

\begin{theorem}[Corollary to Sturm's Theorem]
  \label{thm:Sturm}
  \uses{def:real_closure}
  Let $F$ be an ordered field, and let $f$ be a polynomial over $F$. Then $f$ has the same number of roots in any real closure of $F$.
\end{theorem}
\begin{proof}
  TODO : decide on the generality of the statement of Sturm's Theorem
\end{proof}


\begin{lemma}
  \label{lem:closure_emb_ext_unordered}
  \uses{def:real_closure}
  Let $F$ be an ordered field with a real closure $R$, and let $K/F$ be a finite ordered extension. Then there is an $F$-homomorphism $K\to R$.
\end{lemma}
\begin{proof}
  \uses{lem:real_closure_exists,thm:Sturm}
  By the primitive element theorem, $K=F(\alpha)$ for some $\alpha\in K$. Let $f$ be the minimal polynomial of $\alpha$ over $F$. Since $F$ has a root in $K$, it has a root in a real closure of $K$ (one exists by Lemma \ref{lem:real_closure_exists}). By Theorem \ref{thm:Sturm}, $f$ has a root $\beta$ in $R$. Therefore define $\varphi:K\to R$ with $\varphi(\alpha)=\beta$.
\end{proof}

\begin{lemma}
  \label{lem:closure_emb_ext}
  \uses{def:real_closure}
  Let $F$ be an ordered field with a real closure $R$, and let $K/F$ be a finite ordered extension. Then there is a unique order-preserving $F$-homomorphism $K\to R$.
\end{lemma}
\begin{proof}
  \uses{lem:real_closure_exists,thm:Sturm,lem:closure_emb_ext_unordered}
  Fix a real closure $R'$ of $K$ (one exists by Lemma \ref{lem:real_closure_exists}).
  
  By the primitive element theorem, $K=F(\alpha)$ for some $\alpha\in K$. Let $f$ be the minimal polynomial of $\alpha$ over $F$, and let $\alpha_1<\cdots<\alpha_m$ be the roots of $f$ in $R'$, with $\alpha=\alpha_k$. By Theorem \ref{thm:Sturm}, $f$ also has $m$ roots in $R$; let them be $\beta_1<\cdots<\beta_m$. Since non-negative elements of $R'$ are squares, and $x_1,\dots,x_{m-1}\in R'$ such that $\alpha_{j+1}-\alpha_j=x_j^2$, and let $L=K(\alpha_1,\dots,\alpha_m,r,x_1,\dots,x_{m-1})\leq R'$. Now, suppose we have a $K$-homomorphism $\psi:L\to R$. Each $\psi(\alpha_j)$ is equal to a different $\beta_i$. Then $\psi(\alpha_{j+1})-\psi(\alpha_j)=\psi(x_j)^2\geq0$, so $\psi(\alpha_1)<\cdots<\psi(\alpha_m)$, and so $\psi(\alpha_j)=\beta_j$ for all $j$.

  By Lemma \ref{lem:closure_emb_ext_unordered}, there is in fact an $F$-homomorphism $\varphi:L\to R$. We will show that $\varphi$ is order-preserving. Indeed, fix a non-negative element $x\in L$. As before, find $r\in R'$ such that $x=r^2$, and let $M=L(r)\leq R'$. Apply Lemma \ref{lem:closure_emb_ext_unordered} again to obtain an $L$-homomorphism $\psi:M\to R$; then $\varphi(x)=\psi(x)=\psi(r)^2\geq0$. Therefore $\varphi$ maps non-negative elements to non-negative elements, and so is order-preserving. Then $\varphi|_K$ is the map we want. Note that $\varphi(\alpha)=\beta_k$.

  To see uniqueness, let $\tilde\varphi:K\to R$ be an order-preserving $F$-homomorphism; by existence, $\tilde\varphi$ extends to a an order-preserving $K$-homomorphism $\tilde\psi:L\to R$. Then $\tilde\varphi(\alpha)=\tilde\psi(\alpha_k)=\beta_k=\varphi(\alpha)$, and so $\tilde\varphi=\varphi$.
\end{proof}

Taking $K=F$ above, we see that the order-embedding of a field into its real closure is unique.

\begin{theorem}
  \label{thm:real_closure_unique}
  \uses{def:real_closure}
  Let $F$ be an ordered field. Then the real closure of $F$ is unique up to unique $F$-isomorphism.
\end{theorem}
\begin{proof}
  \uses{lem:closure_emb_ext,lem:RCF_max}
  Let $R_1$ and $R_2$ be real closures of $F$. Applying Zorn's lemma to the set of ordered extensions intermediate between $R_1$ and $F$ having a unique order-preserving $F$-embedding into $R_2$, and using Lemma \ref{lem:closure_emb_ext}, we obtain an intermediate extension $R_1/K/F$ with no nontrivial finite ordered extensions and a unique order-preserving $F$-embedding $\varphi:K\to R_2$. If the ordered algebraic extension $R_2/\varphi(K)$ were nontrivial, then it would contain a nontrivial ordered finite extension, so $\varphi$ must be surjective (and so an $F$-isomorphism). In particular, $K\subseteq R_1$ is real closed; by maximality (Lemma \ref{lem:RCF_max}), in fact $K=R_1$ and so $\varphi$ is an $F$-isomorphism between $R_1$ and $R_2$.
\end{proof}

\begin{corollary}
  \label{cor:RCF_no_auto}
  \uses{def:RCF}
  A real closed field has no nontrivial field automorphisms.
\end{corollary}
\begin{proof}
  \uses{thm:real_closure_unique, lem:RCF_ord_unique}
  Let $R$ be a real closed field. By Theorem \ref{thm:real_closure_unique}, $R$ has no nontrivial order-preserving automorphisms. Since the ordering on $R$ is unique (by Lemma \ref{lem:RCF_ord_unique}), every automorphism of $R$ must be order-preserving.
\end{proof}

This uniqueness result is stronger than the one in the algebraically closed case: an algebraically closed field has many nontrivial automorphisms.

Uniqueness of algebraic closures allows us to classify ordered algebraic extensions.

\begin{lemma}
  \label{lem:ord_alg_ext_eq_emb}
  \uses{def:real_closure}
  Let $F$ be an ordered field with real closure $R$, and let $K/F$ be algebraic. Then field orderings on $K$ making $K/F$ ordered correspond to $F$-homomorphisms $K\to R$ via the order obtained by restriction from $R$.
\end{lemma}
\begin{proof}
  \uses{lem:real_closure_exists,thm:real_closure_unique}
  Fix an ordering on $K$ extending that on $F$, and let $K$ have real closure $R_K$ (exists by Lemma \ref{lem:real_closure_exists}). Then $R_K/F$ is algebraic, so $R_K$ is a real closure of $F$. By Theorem \ref{thm:real_closure_unique}, there is an $F$-isomorphism to $R_K\cong R$, and this induces an $F$-homomorphism $K\to R$. Restricting the order on $R$ to $K$ via this map recovers the original order on $K$ by construction.
  
  Moreover, the inverse to order restriction constructed above is unique. Indeed, an inverse $\varphi:K\to R$ is order-preserving by definition, so it is an order-embedding from $K$ into its real closure. By Theorem \ref{thm:real_closure_unique}, such a map is unique.
\end{proof}

\section{The Artin-Schreier Theorem}

We can go much further than Lemma \ref{lem:FTAlg_converse_alg_closure}. The following is a weak form of the Artin-Schreier theorem. Removing the condition on characteristic is possible, but requires some more involved algebra.

\begin{theorem}
  \label{thm:Artin-Schreier_weak}
  \uses{def:RCF}
  \lean{weak_Artin_Schreier}
  \leanok
  Let $R$ be a field with $\ch K\neq2$, and suppose $[\bar{R}:R]=2$. Then $R$ is real closed.
\end{theorem}
\begin{proof}
  \uses{lem:FTAlg_converse_alg_closure}
  By Lemma \ref{lem:FTAlg_converse_alg_closure}, it suffices to show that $\bar{R}\cong R(i)$.

  Since $\ch\bar{R}\neq 2$, we have $\bar{R}\cong R(\sqrt{a})$ for some non-square $a\in R$. Since $R(\sqrt{a})$ is algebraically closed, $\sqrt{a}$ is a square in $R(\sqrt{a})$; find $x,y\in R$ such that $\sqrt{a}=(x+y\sqrt{a})^2$. Expanding and comparing coefficients, $x^2+y^2a=0$ and $2xy=1$. Rearranging, $a=-(4x^4)=-1\cdot(2x^2)^2$. Rescaling by $2x^2$, $R(i)\cong R(\sqrt{a})=\bar{R}$.
\end{proof}

We can weaken the hypotheses even further. Here is the full Artin-Schreier theorem.

\begin{theorem}[Artin-Schreier Theorem]
  \label{thm:Artin-Schreier}
  \uses{def:RCF}
  \lean{Artin_Schreier}
  \leanok
  Let $R$ be a field, and suppose $\bar{R}$ is a \textit{finite} extension of $R$. Then $R$ is real closed.
\end{theorem}
\begin{proof}
  \uses{thm:Artin-Schreier_weak}
  TODO: this needs a lot more preliminaries eg Artin-Schreier theory, Kummer theory
\end{proof}

\begin{corollary}
  \label{cor:ACF_char_p_no_fin_ind}
  An algebraically closed field of nonzero characteristic has no finite-index subfields.
\end{corollary}
\begin{proof}
  \uses{thm:Artin-Schreier}
  Ordered fields have characteristic 0.
\end{proof}

\begin{theorem}
  \label{thm:Q_alg_unique_fin_ind}
  The finite-index subfields of $\bar{\Q}$ are isomorphic copies of $\Q_{\text{alg}}=\bar{\Q}\cap\R$ indexed by $\Gal(\Q_{\text{alg}}/\Q(i))$.
\end{theorem}
\begin{proof}
  \uses{thm:Artin-Schreier,thm:real_closure_unique,cor:unique_ord_Q}
  Since $\Q_{\text{alg}}(i)=\bar{\Q}$, the field $\Q_{\text{alg}}$ is a finite-index subfield.

  By Theorem \ref{thm:Artin-Schreier}, any finite-index subfield has a unique ordering making it real closed. Let $R$ be a real closed subfield. Then the ordering on $R$ restricts to the ordering on $\Q$ (unique by Corollary \ref{cor:unique_ord_Q}). Since $R/\Q$ is algebraic, $R$ is a real closure of $\Q$. By Theorem \ref{thm:real_closure_unique}, there is a unique $\Q$-isomorphism $\psi:\Q_{\text{alg}}\cong R$. Since $\Q_{\text{alg}}(i)=R(i)=\bar{\Q}$, $\psi$ extends uniquely to $\varphi\in\Gal(\Q_{alg}/\Q(i))$ by mapping $i\to i$. The subfield $R$ is then recovered from $\varphi\in\Gal(\Q_{alg}/\Q(i))$ by taking $\varphi^{-1}(\Q_{\text{alg}})$.
\end{proof}

\bibliographystyle{plain}
\bibliography{rcf}